\begin{centering}
\textbf{ACKNOWLEDGEMENTS}\\
\vspace{\baselineskip}
\end{centering}

First and foremost, I would like to thank my advisor, Prof. Jeremy Siek for his
outstanding guidance and unwavering support. Prof. Siek is my role model for
being a researcher. He encourages me to think in a deep and rigorous way while
presenting the results in a lucid and accessible way. Prof. Siek is also my role
model for being an educator. He shows me how to present the course materials in
an incremental and easily accessible way that is relatable to the experiences of
the students. I owe all the achievements during my graduate studies to Prof.
Siek.

I would like to thank the professors in my Ph.D. advisory committee (in no
particular order): Prof. Amr Sabry, Prof. Chung-chieh Shan, and Prof. Sam
Tobin-Hochstadt for providing feedback and suggestions to my research and this
dissertation. In addition, I would like to thank Prof. Dan Friedman: like many
students at Indiana University, I was introduced to the world of programming
languages by Prof. Friedman's ``little books'' and the lectures of C311/B521
``Programming Language Principles.''

I would like to thank the exceptional community of programming language
researchers at Indiana University. I am especially grateful to the following
professors, students, and postdoctoral researchers (in no particular order):
Prof. Carlo Angiuli, Prof. Ryan Newton, Dr. David Christiansen, Joshua Crotts,
Caner Derici, Chenchao Ding, Fred Fu, Aria Givens, Ethan Hawk, Artem Iurchenko,
Sanad Kadu, Caleb Schultz Kisby, Darshal Shetty, Zixiu Su, Jifeng Wu, Yafei
Yang, Andre Kuhlenschmidt, Ryan Scott, Tulip Amalie, Kartik Sabharwal, Annie
Pompa, Vikraman Choudhury, Sam Bowman, Chaitanya Koparkar, Rajan Walia, Sarah
Spall, Weixi Ma, Matthew Heimerdinger, Chao-Hong Chen, Michael Vollmer (now
professor at the University of Kent), Victoria Vollmer, Kuang-Chen Lu, Joshua
Larkin, Deyaaeldeen Almahallawi, Aaron Hsu, Paulette Koronkevich, Andrew Kent,
and Jason Hemann (now professor at Seton Hall University). \textit{Thank you!}

I thank Zihan Chen, Che Jingyin, Katharine Khamhaengwong, Jacob Striebel,
Xiaorui Pan, and Kan Yuan at Indiana University for friendship.

I am grateful to all the students whom I taught. In particular, I would like to
thank Calvin Josenhans, Gautam Hari, Sparsh Nair, Oleksandra Tkachuk, Zeshawn
Zahid, Jacob Herbert, Dylan Jacoby, Solomon Zinn Krulewitch, and Yuntian Zeng. I
also thank Dr. Akesha Horton for her lectures on pedagogy.

I thank Dr. Vilhelm Sj\"{o}berg for hosting me as a summer intern on software
verification at CertiK. I thank Prof. Kristopher Micinski, Prof. Sergey Bratus,
and Prof. Gang Tan for advice and discussion.

I would like to thank the operating system research group at Columbia University
led by Prof. Junfeng Yang, who provided a valuable summer research opportunity
to me when I was an undergrad. I appreciate the friendship and help from (in no
particular order) Dr. Heming Cui (now professor at the University of Hong Kong),
Rui Gu, Yang Tang (now professor at New York University), Gang Hu, Xinhao Yuan,
Lingmei Weng, and Chang Lou (now professor at the University of Virginia). I
thank Prof. Yu-Ping Wang for lecturing me on symbolic execution while I was
doing undergraduate independent study at Tsinghua. Also, I express special
gratitude to the friends and roommates from my undergraduate days at Tsinghua:
Hongyin Luo, Yuan Yang, Hao Wang, Hongyi Wen (now professor at New York
University Shanghai), Xintian Li, Hanxuan Yu, and S\^{e}kai Zhou.

I would like to thank Hattie Fu for her love, support, and company during the
final (and one of the hardest) stage of my dissertation writing.

I thank my grandmother, who taught physics and electrical engineering before
retirement, for introducing me to mathematics and applied sciences. I thank my
aunts for buying me the first computer in my life: a used graphic workstation
with an Intel 80486 processor. Lastly, and most importantly, I would like to say
thank you to my parents. I could not have completed this quest without your
support and understanding.

(This dissertation is based upon work supported by the National Science
Foundation under Grant No. 1763922 ``Performant Sound Gradual Typing.'')
